How to select predictive models for decision-making or causal inference?

### 2024-09-11 how_to_select_latex.tex
apd:simus add generation algo

% added commands and packages
\newtheorem{proof}{Proof}
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
\newtheorem{proposition}[theorem]{Proposition}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{proposition*}{Proposition}
\newcommand{\indep}{\perp \!\!\! \perp}
\newcommand{\eg}{\emph{e.g.}}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm, algorithmicx, algpseudocode}
\DeclareMathOperator*{\argmin}{argmin} \def\mycitecolor{green!50!black}
\usepackage{mathtools}
\newcommand\myeq{\stackrel{\mathclap{\text{def}}}{=}}
\usepackage{booktabs} % For formal tables (eg. bottomrules command)
\usepackage{multirow} % For multirow in tables
\usepackage{threeparttable} % For table 
\usepackage{caption} % for subfigures
\usepackage{subcaption}
\usepackage{placeins} % for FloatBarrier command to prevent a figure to pass some point.
\usepackage{titlesec} % for title formatting
%\titleformat*{\subsection}{\color{jnlclr}\bfseries} % force colored subsection font for paragraphs
%\titleformat*{\paragraph}{\bfseries} % force bold font for paragraphs

@@

We use Gaussian-distributed covariates and random basis expansion based on
Radial Basis Function kernels. A random basis of RBF kernel enables modeling
non-linear and complex relationships between covariates in a similar way to the
well known spline expansion. The estimators of the response function are learned
with a linear model on another random basis (which can be seen as a stochastic
approximation of the full data kernel \cite{rahimi_random_2008}). We carefully
control the amount of overlap between treated and control populations, a crucial
assumption for causal inference. Figure \ref{fig:simulation_examples}
illustrates 2D examples of the simulation.

@@

    \item Functions $\mu_0$, $\tau$ are distinct linear functions of the
          transformed features:
          \begin{equation*}
              \mu_0(x) = \begin{bmatrix} z(x); 1 \end{bmatrix} \cdot \beta_{\mu}^T
          \end{equation*}
          \begin{equation*}
              \tau(x) = \begin{bmatrix} z(x); 1 \end{bmatrix} \cdot \beta_{\tau}^T
          \end{equation*}
          Where $\beta_{\mu}$ and $\beta_{\tau}$ are sampled from two normal distributions with mean 0 and unit variances $\mathcal{N}(0, I_{D+1})$.
    \item Adding a Gaussian noise, $\varepsilon \sim \mathcal N(0, \sigma(x;a))$,
          we construct the potential outcomes:
          $y(a) = \mu_0(x) + a\,\tau(x) + \varepsilon(x, a)$
\end{itemize}
We generated 1000 instances of this dataset with uniformly random overlap
parameters $\theta \in \left[ 0, 2.5 \right]$.

>>>>

% added commands and packages
\newtheorem{proof}{Proof}
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
\newtheorem{proposition}[theorem]{Proposition}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{proposition*}{Proposition}
\newcommand{\indep}{\perp \!\!\! \perp}
\newcommand{\eg}{\emph{e.g.}}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm, algorithmicx, algpseudocode}
\DeclareMathOperator*{\argmin}{argmin} \def\mycitecolor{green!50!black}
\usepackage{mathtools}
\newcommand\myeq{\stackrel{\mathclap{\text{def}}}{=}}
\usepackage{booktabs} % For formal tables (eg. bottomrules command)
\usepackage{multirow} % For multirow in tables
\usepackage{threeparttable} % For table 
\usepackage{caption} % for subfigures
\usepackage{subcaption}
\usepackage{placeins} % for FloatBarrier command to prevent a figure to pass some point.
\usepackage{titlesec} % for title formatting
%\titleformat*{\subsection}{\color{jnlclr}\bfseries} % force colored subsection font for paragraphs
%\titleformat*{\paragraph}{\bfseries} % force bold font for paragraphs
\usepackage{algpseudocode} % for algorithmic

@@

We use Gaussian-distributed covariates and random basis expansion based on
Radial Basis Function kernels. A random basis of RBF kernel enables modeling
non-linear and complex relationships between covariates in a similar way to the
well known spline expansion. The estimators of the response function are learned
with a linear model on another random basis (which can be seen as a stochastic
approximation of the full data kernel \cite{rahimi_random_2008}). We carefully
control the amount of overlap between treated and control populations, a crucial
assumption for causal inference. Algorithm \ref{apd:experiments:generation:algo} describes the generation process for one simulation. Figure \ref{fig:simulation_examples}
illustrates 2D examples of the simulation.

@@

    \item Functions $\mu_0$, $\tau$ are distinct linear functions of the
          transformed features:
          \begin{equation*}
              \mu_0(x) = \begin{bmatrix} z(x); 1 \end{bmatrix} \cdot \beta_{\mu}^T
          \end{equation*}
          \begin{equation*}
              \tau(x) = \begin{bmatrix} z(x); 1 \end{bmatrix} \cdot \beta_{\tau}^T
          \end{equation*}
          Where $\beta_{\mu}$ and $\beta_{\tau}$ are sampled from two normal distributions with mean 0 and unit variances $\mathcal{N}(0, I_{D+1})$.
    \item Adding a Gaussian noise, $\varepsilon \sim \mathcal N(0, \sigma(x;a))$,
          we construct the potential outcomes:
          $y(a) = (1-\omega)\mu_0(x) + a\,\omega\tau(x) + \varepsilon(x, a)$ with $\omega$ the treatment effect relative size with respect to the baseline response.
\end{itemize}
We generated 1000 instances of this dataset with uniformly random overlap
parameters $\theta \in \left[ 0, 2.5 \right]$.

\begin{algorithm}
    \caption{Data simulation for the simulated dataset caussim}
    \label{apd:experiments:generation:algo}
    Let the parameters of a given simulation be:
    \begin{itemize}
        \item N the number of samples.
        \item D the dimension of the basis expansion.
        \item $p_A$ the proportion of treated individuals.
        \item $\theta$ the overlap parameter.
        \item $\mathbb P(x|A=a) = W \cdot \mathcal N \Big( \begin{bmatrix} (1-2a) \theta \\ 0\end{bmatrix} ; \begin{bmatrix} \sigma_0 & 0 \\ 0 & \sigma_1\end{bmatrix} \Big)$ the rotated gaussian for each population covariate. $W$ is a random rotation matrix and $\sigma_0^2=2;\sigma_1^2=5$ are scaling parameters.
        \item $e(x)=\mathbb P(A=1|X=x)$ the oracle propensity score, obtained analytically from $\mathbb P(x|A=a)$.
        \item $\mathbb P(X) = p_A \mathbb P(X|A=1) + (1- p_A) \mathbb P(X|A=0)$ the mixture of gaussian for the whole population covariates.
        \item $b_1, .., b_D \sim \mathbb P(X)$, the basis sampled from the gaussian mixture.
        \item $z(x) = [RBF_{\gamma}(x, b_d)]_{d=1..D} \cdot Z^T \in
                  \mathbb{R}^D$ the Nystroem expansion. $RBF_{\gamma}$ is the
              Gaussian kernel $K(x, y) = exp(-\gamma ||x-y||^2)$ and Z is the
              normalization constant of the kernel basis, computed as the root
              inverse of the basis kernel $Z=[K(b_i, b_j)]_{i, j \in
                  {1..D}}^{-1/2}$.
        \item $\beta_{\mu}$ and $\beta_{\tau}$, the linear coefficients on top of the basis expansion. They are sampled from two normal distributions with mean 0 and unit variances $\mathcal{N}(0, I_{D+1})$.
        \item $\omega$ the treatment effect relative size with respect to the baseline response.
        \item $\varepsilon(x, a) \sim \mathcal N(0, \sigma_y)$ an exogen gaussian noise with $\sigma_y$ the scale of the noise.
    \end{itemize}

    Generation process for one sample ($x$, $a$, e(x), $y(0)$, $y(1)$, $y$):
    \begin{algorithmic}[1]
        \State $x \sim \mathbb P(X)$ \Comment{Sample from rotated gaussian mixture}
        \State  $z(x)$ \Comment{Compute the Nystroem expansion}
        \State $\mu_0(x) = \begin{bmatrix} z(x); 1 \end{bmatrix} \cdot \beta_{\mu}^T$ \Comment{Compute the non-treated response surface, ie. the baseline}
        \State  $\tau(x) = \begin{bmatrix} z(x); 1 \end{bmatrix} \cdot \beta_{\tau}^T$ \Comment{Compute the conditional treatment effect (CATE)}
        \State $y(a) = (1-\omega) \mu_0(x) + a\,\omega\tau(x) + \varepsilon(x, a) + \varepsilon(x, a)$ \Comment{Compute the potential outcomes}
        \State $a \sim \operatorname{Bernoulli}(e(x))$ \Comment{Compute the treatment status}
        \State $y = a\,y(1) + (1-a)\,y(0)$ \Comment{Compute the observed outcome}
    \end{algorithmic}

\end{algorithm}
